Investigating the use of aesthetic measures over time with unsupervised evolutionary art

### 2021-03-29 Paper/paper.tex
Added godot showing automation and also shannon entropy

\section*{Evolutionary Algorithm}
The evolution of shaders is done by our own software. In the program, shaders are represented through 3 expression trees, with one for red, blue, and green colors. In these trees, there are 
operator nodes, which have two children, and leaf nodes which do not have children. All operators have 2 children to avoid making invalid trees during mutation and crossover stages. 
Operators are split into 3 types, which determine how their children are used in calculation and writing. Type 1 operators use a "left right" orientation, for example, the plus operator: Uv.x + 3. 
Type 2 operators are functions with 2 inputs, for example the noise operator, noise(2,4). Type 3 operators are functions with only one input, and to preserve each operator having 2 children, for example the PlusCos operator: 3 + cos(14).
\\
\begin{tabular}{|c|c|c|c|}
    \hline
    Operator & Type&  Example & Description\\
    \hline
    Plus & 1 & Uv.x + 32 & Standard Addition \\
    \hline
    Minus & 1 &TIME - 3 & Standard Subtraction \\
    \hline
    Multiply & 1 & 15 * Uv.y & Standard Multiplication  \\
    \hline
    Divide & 1 & 23 /32 & Standard Division  \\
    \hline
    PlusCos & 3 & Uv.y Cos(34) & Left child added to cosine of right child  \\
    \hline
    MultCos & 3 & 20 * Cos(Uv.x) & Left child multiplied by cosine of right child  \\
    \hline
    PlusSin & 3  & 24 + sin(32) & Left child added to sin of right child  \\
    \hline
    Multiply & 3 & 15 * sin(21)& Left child multiplied by sine of right child   \\
    \hline
\end{tabular}\\\\
Leaf nodes are further split into constants, and variables. 
\\\\
\begin{tabular}{|c|c|}
    \hline
    Variable & Description \\ 
    \hline
    Uv.x & x coordinate of current pixel, normalized [0,1]\\
    \hline
    Uv.y & y coordinate of current pixel, normalized [0,1]\\
    \hline
    TIME & Time since shader started\\
    \hline
\end{tabular}\\\\
These expression trees are traversed inorder to write the trees into a Godot shader. The shader's trees are solved to produce a matrix of colors at different time steps. The average difference between every timestep and its following timestep is then computed,
and then the average of those averages is added to the fitness score, with shdaders that vary more being favored over more static shaders.
\end{document}

>>>>

\section*{Evolutionary Algorithm}
\subsection*{Representation of shaders}
The evolution of shaders is done by our own software. In the program, shaders are represented through 3 expression trees, with one for red, blue, and green colors. In these trees, there are 
operator nodes, which have two children, and leaf nodes which do not have children. All operators have 2 children to avoid making invalid trees during mutation and crossover stages. 
Operators are split into 3 types, which determine how their children are used in calculation and writing. Type 1 operators use a "left right" orientation, for example, the plus operator: Uv.x + 3. 
Type 2 operators are functions with 2 inputs, for example the noise operator, noise(2,4). Type 3 operators are functions with only one input, and to preserve each operator having 2 children, for example the PlusCos operator: 3 + cos(14).
\\
\begin{tabular}{|c|c|c|c|}
    \hline
    Operator & Type&  Example & Description\\
    \hline
    Plus & 1 & Uv.x + 32 & Standard Addition \\
    \hline
    Minus & 1 &TIME - 3 & Standard Subtraction \\
    \hline
    Multiply & 1 & 15 * Uv.y & Standard Multiplication  \\
    \hline
    Divide & 1 & 23 /32 & Standard Division  \\
    \hline
    PlusCos & 3 & Uv.y Cos(34) & Left child added to cosine of right child  \\
    \hline
    MultCos & 3 & 20 * Cos(Uv.x) & Left child multiplied by cosine of right child  \\
    \hline
    PlusSin & 3  & 24 + sin(32) & Left child added to sin of right child  \\
    \hline
    Multiply & 3 & 15 * sin(21)& Left child multiplied by sine of right child   \\
    \hline
\end{tabular}\\\\
Leaf nodes are further split into constants, and variables. 
\\\\
\begin{tabular}{|c|c|}
    \hline
    Variable & Description \\ 
    \hline
    Uv.x & x coordinate of current pixel, normalized [0,1]\\
    \hline
    Uv.y & y coordinate of current pixel, normalized [0,1]\\
    \hline
    TIME & Time since shader started\\
    \hline
\end{tabular}\\\\
These expression trees are traversed inorder to write the trees into a Godot shader. The shader's trees are solved to produce a matrix of colors at different time steps. The average difference between every timestep and its following timestep is then computed,
and then the average of those averages is added to the fitness score, with shdaders that vary more being favored over more static shaders.
\subsection*{Evolving shaders}
To start, the algorithm is initialized by shaders with random numbers of nodes and randomized values. Then each shader has its fitness score assigned, by the average difference over time and the chosen aesthetic measure. 
After the shaders have had their fitness value calculated X number of survivors are selected. The survivors are then paired up, and crossed over by swapping 2 nodes (and by extension its children) on each red, blue, and green
tree. Then Y copies are made of each survivor, and then mutated. Then the procedure is repeated until desired stopping point.
\end{document}
